Visualising gene differences based on read alignments

### 2022-03-08 analysis/GeneLossComparisonReads.Rmd
workflowr::wflow_publish(files = "analysis/GeneLossComparisonReads.Rmd")

```{r}
stomata_long <- stomata %>% pivot_longer(-Gene.symbol, values_to = 'Coverage', names_to='Species') %>% 
  mutate(Species = gsub('\\.\\.', '. ', Species))
```

@@

@@

```{r}
ethylene_long <- ethylene %>% pivot_longer(-Gene.symbol, values_to = 'Coverage', names_to='Species') %>% 
  mutate(Species = gsub('\\.\\.', '. ', Species))
```

@@

  theme(legend.position = "None",
  axis.text.y = element_text(face = "italic")) +
  scale_color_brewer(palette='Dark2') +
  ylab('Coverage (%)') +
  coord_flip() +
  xlab('Species')
eth_p
```
# NDH genes

@@

```{r}
ndh_long <- ndh %>% pivot_longer(-Gene.symbol, values_to = 'Coverage', names_to='Species') %>% 
  mutate(Species = gsub('\\.\\.', '. ', Species))
```

@@

  theme(legend.position = "None",
  axis.text.y = element_text(face = "italic")) +
  scale_color_brewer(palette='Dark2') +
  ylab('Coverage %') +
  coord_flip() +
  xlab('Species')
ndh_p
```

@@

```{r}
ndh_chl_long <- ndh_chloro %>% pivot_longer(-Gene.symbol, values_to = 'Coverage', names_to='Species') %>% 
  mutate(Species = gsub('\\.\\.', '. ', Species))
```

@@

@@

>>>>

```{r}
stomata_long <- stomata %>% pivot_longer(-Gene.symbol, values_to = 'Coverage', names_to='Species') %>% 
  mutate(Species = gsub('\\.\\.', '. ', Species)) %>% 
  filter(Species != 'H. ovalis')
```

@@

```{r}
label <- stomata_long #%>% mutate(Gene.symbol = stringr::str_to_lower(Gene.symbol)) 
label <- label %>% 
  mutate(mylabel = case_when(Coverage >= 100 ~ '',
                             Coverage == 0 ~ '',
                                    TRUE ~ as.character(Coverage)))

stom_heat <- stomata_long %>% 
  #mutate(Gene.symbol = stringr::str_to_lower(Gene.symbol)) %>% 
  ggplot(aes(x=Species, y=factor(Gene.symbol, level=rev(unique(Gene.symbol))), fill=Coverage)) + 
  geom_tile(alpha=0.9, color='white') + 
  geom_text(aes(label=mylabel), data=label, size=3.5) + 
  theme(axis.text.y = element_text(face = "italic"),
        axis.text.x = element_text(face='italic'),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none") +
  scale_fill_continuous(limits=c(0,100))

stom_heat
```

@@

```{r}
ethylene_long <- ethylene %>% pivot_longer(-Gene.symbol, values_to = 'Coverage', names_to='Species') %>% 
  mutate(Species = gsub('\\.\\.', '. ', Species)) %>% 
  filter(Species != 'H. ovalis')
```

@@

  theme(legend.position = "None",
  axis.text.y = element_text(face = "italic")) +
  scale_color_brewer(palette='Dark2') +
  ylab('Coverage (%)') +
  coord_flip() +
  xlab('Species')
eth_p
```

```{r}
eth_heat <- ethylene_long %>% 
  dplyr::filter(Gene.symbol != 'ndhB-2') %>% 
  mutate(Gene.symbol = stringr::str_to_lower(Gene.symbol)) %>% 
  ggplot(aes(x=Species, y=factor(Gene.symbol, level=rev(unique(Gene.symbol))), fill=Coverage)) + 
  geom_tile() + 
  ylab('Gene') +
  theme(axis.text.y = element_text(face = "italic"),
        axis.text.x = element_text(face='italic')) +
  scale_fill_continuous(limits=c(0,100))

eth_heat
```

# NDH genes

@@

```{r}
ndh_long <- ndh %>% pivot_longer(-Gene.symbol, values_to = 'Coverage', names_to='Species') %>% 
  mutate(Species = gsub('\\.\\.', '. ', Species)) %>% 
  filter(Species != 'H. ovalis')
```

@@

  theme(legend.position = "None",
  axis.text.y = element_text(face = "italic")) +
  scale_color_brewer(palette='Dark2') +
  ylab('Coverage %') +
  coord_flip() +
  xlab('Species')
ndh_p
```
```{r fig.width=10}
label <- ndh_long
label <- label %>% 
  mutate(mylabel = case_when(Coverage >= 100 ~ '',
                             Coverage == 0 ~ '',
                                    TRUE ~ as.character(Coverage)))

ndh_heat <- ndh_long %>% 
  dplyr::filter(Gene.symbol != 'ndhB-2') %>% 
  mutate(Gene.symbol = stringr::str_to_lower(Gene.symbol)) %>% 
  ggplot(aes(x=Species, y=factor(Gene.symbol, level=rev(unique(Gene.symbol))), fill=Coverage)) + 
  geom_tile() + 
  geom_tile(alpha=0.9, color='white') + 
  #geom_text(data=label, aes(label=mylabel), size=3.5) +
  theme(axis.text.y = element_text(face = "italic"),
        axis.text.x = element_text(face = 'italic'),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())  +
  scale_fill_continuous(limits=c(0,100))

ndh_heat
```

@@

```{r}
ndh_chl_long <- ndh_chloro %>% pivot_longer(-Gene.symbol, values_to = 'Coverage', names_to='Species') %>% 
  mutate(Species = gsub('\\.\\.', '. ', Species)) %>% 
  filter(Species != 'H. ovalis')
```

@@

```{r}
label <- ndh_chl_long %>% filter(Gene.symbol != 'ndhB-2')
label <- label %>% 
  mutate(mylabel = case_when(Coverage >= 100 ~ '',
                             Coverage == 0 ~ '',
                                    TRUE ~ as.character(Coverage)))

ndh_chl_heat <- ndh_chl_long %>% 
  dplyr::filter(Gene.symbol != 'ndhB-2') %>% 
  mutate(Coverage = case_when(Coverage > 100 ~ 100,
                              TRUE ~ Coverage)) %>% 
  ggplot(aes(x=Species, y=factor(Gene.symbol, level=rev(unique(Gene.symbol))), fill=Coverage)) + 
  geom_tile(alpha=0.9, color='white') + 
  geom_text(data=label, aes(label=mylabel), size=3.5) +
  ylab('Gene') +
  theme(axis.text.y = element_text(face = "italic"),
        axis.text.x = element_text(face = 'italic'),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none") +
  scale_fill_continuous(limits=c(0,100))

ndh_chl_heat
```

@@

```{r fig.height=7}
patchwork <- (eth_p + ndh_chl_heat)/
 (ndh_p + stom_heat  )
#patchwork[[2]] = patchwork[[2]] + theme(axis.title.x = element_blank() )
#patchwork[[1]] = patchwork[[1]] + theme(axis.title.x = element_blank() )
puh <- patchwork+ plot_annotation(tag_levels = "A") #+ plot_layout(guides='none')
puh
cowplot::save_plot(filename = 'output/patched_gene_loss.png', plot = puh, base_height = 7)
```

Hmmmmmmmmm
